Illuminating a Dark Lens : A Type Ia Supernova \\ Magnified by Galaxy Cluster Abell 2744

### 2015-05-01 snTomas.tex
really minor tweaks to text. Added some caveats on unblind models

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
\change{$z=1.3457\pm0.0001$} behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
extinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.77\pm0.15$} magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.03\pm0.29$}.  We compare this direct
measurement of the magnification against predicted magnifications
from \changeR{15 lens models developed by 8 independent groups},
representing a broad range of methodologies and incorporating a
variety of strong- and weak-lensing constraints.  The models are
collectively fairly accurate: 8 of the models deliver median
magnifications that are consistent with the measured $\mu$ to within
the 1$\sigma$ uncertainties.  \changeR{However, the models exhibit a
subtle systematic bias: all of the models that disagree with the SN by
more than $1\sigma$ are {\it higher} than the measured value}
and \change{6 of these models are $\geq1.5\sigma$} above the measured
magnification.  We evaluate possible causes for this mild
bias, \changeR{and reject explanations that invoke modeling errors or
misinterpretation of the SN data. We conclude that -- at least for
this cluster and this line of sight -- the most accurate models are
those that employ stringent quality cuts for strong-lensing
constraints.} \change{A more robust evaluation of systematic biases
could be achieved with further cooperative lens model comparisons
using a larger sample of lensed Type Ia SNe as well as cluster lensing
simulations.}
\end{abstract}

@@

\changeR{
However, the right panel of Figure~\ref{fig:Nspecz} shows that 3 of
the most successful lens models (Bradac(v2), Lam(v2), Diego) do share
a common quality: they all use a set of multiply-imaged galaxies for
which a large fraction have spectroscopic redshifts.  In practice,
this means that the strong-lensing constraints in these models are
defined more stringently than other models. Some other models may use
a larger set of strong-lensing features, but the dearth of
spectroscopic constraints means that it is more likely that some of
those multiply-imaged galaxies are assigned with incorrect redshifts.
}

\changeR{
This comparison suggests that to get accurate magnifications, lens
models should employ a strict vetting of the strong-lensing
constraints.  Multiply-imaged systems should ideally have
spectroscopic redshifts, or at least very well constrained photometric
redshifts.  The {\it quantity} of strong-lensing constraints is not as
important as their {\it quality}.
}

>>>>

\begin{abstract}
SN \tomas\ is a Type Ia Supernova (\SNIa) discovered at
\change{$z=1.3457\pm0.0001$} behind the galaxy cluster Abell 2744 ($z=0.308$). This
SN has a projected separation from the cluster core of
$\sim$40\arcsec, closer to the critical lensing line than any previous
cluster-lensed Type Ia SN.  As such, it provides a valuable
opportunity to confront gravitational lens models of galaxy clusters
with the direct measurement of an absolute magnification at the edge
of the strong-lensing region.  We derive a tightly constrained measure
of the peak apparent magnitude, corrected for light curve shape and
extinction.  In a cosmology-independent analysis, we find that \tomas\
is \change{$0.77\pm0.15$} magnitudes brighter than unlensed Type Ia
SNe at similar redshift, implying a lensing magnification
of \change{$\mu_{\rm obs}=2.03\pm0.29$}.  We compare this direct
measurement of the magnification against predicted magnifications
from \changeR{15 lens models developed by 8 independent groups},
representing a broad range of methodologies and incorporating a
variety of strong- and weak-lensing constraints.  The models are
collectively fairly accurate: 8 of the models deliver median
magnifications that are consistent with the measured $\mu$ to within
the 1$\sigma$ uncertainties.  \changeR{However, the models exhibit a
subtle systematic bias: all of the models that disagree with the SN by
more than $1\sigma$ are {\it higher} than the measured value}
and \change{6 of these models are $\geq1.5\sigma$} above the measured
magnification.  We evaluate possible causes for this mild
bias, \changeR{and reject explanations that
invoke} \changeP{misinterpretation of the SN data or errors in the
definition of the lens models}. \changeR{We conclude that -- at least for this
cluster and this line of sight -- the most accurate models are those
that employ stringent quality cuts for strong-lensing
constraints.} \change{ A more robust evaluation of systematic biases
could be achieved with further cooperative lens model comparisons
using a larger sample of lensed Type Ia SNe as well as cluster lensing
simulations.}
\end{abstract}

@@

\changeR{
However, the right panel of Figure~\ref{fig:Nspecz} shows that 3 of
the most successful lens models (Bradac(v2), Lam(v2), Diego) do share
a common characteristic: they all use a set of multiply-imaged galaxies for
which a large fraction have spectroscopic redshifts. In practice,
this means that the strong-lensing constraints in these models are
defined more stringently. Other models use
a larger set of strong-lensing features, but the dearth of
spectroscopic constraints means that it is more likely that some of
those multiply-imaged galaxies are assigned with incorrect redshifts.
}

\changeR{
This comparison suggests that to get accurate magnifications, lens
models should employ a strict vetting of the strong-lensing
constraints.  Multiply-imaged systems should ideally have
spectroscopic redshifts, or at least very well constrained photometric
redshifts.}
\changeP{The {\it quality} of strong-lensing constraints is perhaps just as
important as the {\it quantity}.}

\changeP{
It is relevant to note that the three models highlighted for their
accuracy here (Bradac(v2), Lam(v2), Diego) also happen to be the three
models that are not truly ``blind,'' in that the modelers were aware
of preliminary analysis of the SN magnification.  However, none of
these models used the SN magnification as an input constraint, none
were tuned to match the SN magnification, and the modelers did not
have a final measurement of the SN magnification at the time their
models were generated.}
